---
title: "Homework 4"
author: "Tom Gibbens-Matsuyama"
subtitle: "Prioritizing potential aquaculture"
format: html
editor: visual
execute: 
  eval: true
  warning: false
  message: false
editor_options: 
  chunk_output_type: console
---

## Background
Marine aquaculture can play a major role in the global food supply as a protein source. It has the potential to be more sustainable than land-based meat production. According to Gentry et al. global seafood demand could be met using less than 0.015% of total ocean area. 

The goal of this assignment is to determine potential aquaculture producing areas within the Exclusive Economic Zones (EEZ) on the west coast. A workflow will be produced to find suitable aquaculture locations based on suitable sea surface temperature (SST) and depth values for oysters. This workflow will then be transferred into a function to be applied given certain parameters. 


### Load libraries

```{r}
library(tidyverse)
library(tmap)
library(terra)
library(here)
library(sf)
library(paletteer)
```

### Load data

```{r}
# Shapefile for the West Coast EEZ
eez <- read_sf(here("data", "wc_regions_clean.shp"))

# Depth raster
depth <- rast(here("data", "depth.tif"))

# Load sst data, use file path
files <- list.files(
    here("data"), pattern = "*sst", 
    full.names = TRUE)
  
sst <- c(rast(files))
```

### Compare CRS

```{r}
# Set the all CRSs to EPSG:4326
eez <- st_transform(eez, crs("EPSG:4326"))
depth <- project(depth, crs("EPSG:4326"))
sst <- project(sst, crs("EPSG:4326"))
```


```{r, warning=TRUE}
#| code-fold: true
#| code-summary: "Show the code"

if (st_crs(eez) == st_crs(depth) && st_crs(depth) == st_crs(sst)){
  print("All CRSs are the same")
} else {
  warning("CRSs are NOT the same")
}

```


### Process SST and depth data so they can be combined

```{r}
# Find mean SST from 2008-2012
# Taking the mean of the 5 SST layers and turning it into one layer
mean_sst <- app(sst, fun = mean, na.rm = TRUE)

# Convert average SST from Kelvin to Celsius
sst_c <- mean_sst - 273.15

# Crop depth raster to match extent of SST raster
depth_crop <- crop(depth, sst_c)

# Resample the depth data to match the resolution of the SST data using the nearest neighbor approach
depth_resample <- resample(depth_crop, sst_c, method = "near")

# Check that depth and SST match in resolution, extent and CRS
sst_depth <- c(sst_c, depth_resample)
```

### Find suitable locations
To find suitable locations, we need to reclassify our rasters to places that are and are not suitable for our given species. In this case, oysters can live within a max sea surface  (sst) of 30 degrees Celsius and a min sst of 11 degrees Celsius. Likewise, they ideally live at a max depth of 70 meters up until the surface at 0 meters depth. 
```{r}
# Create reclassification matrix for SST
sst_rcl <- matrix(c(-Inf, 11, 0,
                    11, 30, 1,
                    30, Inf, 0),
                  ncol = 3, byrow = TRUE)

# Use reclassification matrix to reclassify sst raster
sst_reclassify <- classify(sst_c, rcl = sst_rcl)


# Create classification matrix for Depth
depth_rcl <- matrix(c(-Inf, -70, 0,
                      -70, 0, 1,
                      0, Inf, 0),
                    ncol = 3, byrow = TRUE)

# Reclassify Depth raster
depth_reclassify <- classify(depth_resample, rcl = depth_rcl)

# Find areas that satistfy both SST and Depth
# Use lapp() to multiple values of rasters together
sst_depth_condition <- lapp(c(sst_reclassify, depth_reclassify), fun = "*")
```

### Determine the most suitable EEZ

```{r}
# Assign raster values of 0 to NA for cellSize calculation
sst_depth_NA <- sst_depth_condition
sst_depth_NA[sst_depth_NA == 0] <- NA

# Find suitable area
suitable_area <- cellSize(sst_depth_NA,
         mask = TRUE,
         unit = "km")

# Rasterize the eez data
eez_raster <- rasterize(eez, suitable_area, "rgn")

# Sum suitable areas for each eez 
eez_suitable <- zonal(x = suitable_area, 
                      z = eez_raster, 
                      fun = "sum", 
                      na.rm = TRUE)
```


```{r}
#| code-fold: true
#| code-summary: "Show the code"

kableExtra::kable(eez_suitable, col.names = c("Region", "Suitable Area (km\u00B2)"), align = "c")
```


```{r}
#| code-fold: true
#| code-summary: "Show the code"


# Join suitable eez data with the original eez for the map
eez_join <- left_join(eez, eez_suitable, by = "rgn")


# Map of suitable areas for Oysters
tm_shape(depth_crop) +
  tm_raster(palette = paletteer_c("ggthemes::Green-Blue Diverging", 7, direction = -1),
            title = "Depth (m)",
            midpoint = NA) +
  tm_shape(eez_join) +
  tm_polygons(col = "area",
              palette = paletteer_c("ggthemes::Classic Area-Brown", 5),
              title = "Suitable area (km\u00B2)",
              alpha = 0.9) +
  tm_text("rgn", 
          size = 0.45,
          col = "black") + 
  tm_layout(frame = FALSE,
            legend.outside = TRUE,
            legend.text.size = 0.65,
            main.title = "Suitable aquaculture area (km\u00B2) for Oysters along \nthe West Coast Exclusive Economic Zone (EEZ)",
            main.title.size = 1.00) +
  tm_compass(position = c(0, 0.125),
             size = 1.5) +
  tm_scale_bar(position = c(0.05, 0.025)) 
```

### My Function

The California mussel was chosen as the second species because of its high potential to be an aquacultured species. The species is has a depth range of 0 - 100 m and a temperature range of 5 - 22 degrees Celsius. The California mussel can be scene all along the west coast, but it will be interesting to which regions contain the highest habitable area. 
```{r}
#| code-fold: true
#| code-summary: "Show the code"

aqua_fun <- function(species, max_sst, min_sst, max_depth, min_depth ){
  
  # Load eez data
  eez <- read_sf(here("data", "wc_regions_clean.shp"))
  
  # Load depth data
  depth <- rast(here("data", "depth.tif"))
  
  # Load sst data, use file path
  files <- list.files(
    here("data"), pattern = "*sst", 
    full.names = TRUE)
  
  sst <- c(rast(files))
  
  # Set CRS all the same
  eez <- st_transform(eez, crs("EPSG:4326"))
  depth <- project(depth, crs("EPSG:4326"))
  sst <- project(sst, crs("EPSG:4326"))
  
  # Checking that the CRSs reprojected to EPSG:4326
  if (st_crs(eez) == st_crs(depth) && st_crs(depth) == st_crs(sst)){
    print("All CRSs are the same")
  } else {
    warning("CRSs are NOT the same")
  }
  
  # Find mean SST from 2008-2012
  mean_sst <- app(sst, fun = mean, na.rm = TRUE)
  
  # Convert average SST from Kelvin to Celsius
  sst_c <- mean_sst - 273.15
  
  # Crop depth raster to match extent of SST raster
  depth_crop <- crop(depth, sst_c)
  
  # The resolutions of the SST and depth data do not match
  # Resample the depth data to match the resolution of the SST data using the nearest neighbor approach
  depth_resample <- resample(depth_crop, sst_c, method = "near")
  
  
  # Check that depth and SST match in resolution, extent and CRS
  sst_depth <- c(sst_c, depth_resample)
  
  
  
  # Create reclassification matrix for SST
  sst_rcl <- matrix(c(-Inf, min_sst, 0,
                      min_sst, max_sst, 1,
                      max_sst, Inf, 0),
                    ncol = 3, byrow = TRUE)
  
  # Use reclassification matrix to reclassify sst raster
  sst_reclassify <- classify(sst_c, rcl = sst_rcl)

  # Create classification matrix for Depth
  depth_rcl <- matrix(c(-Inf, -max_depth, 0,
                        -max_depth, -min_depth, 1,
                        -min_depth, Inf, 0),
                      ncol = 3, byrow = TRUE)
  
  # Reclassify Depth raster
  depth_reclassify <- classify(depth_resample, rcl = depth_rcl)
  
  # Find areas that satistfy both SST and Depth
  # Use lapp() to multiple values of rasters together
  sst_depth_condition <- lapp(c(sst_reclassify, depth_reclassify), fun = "*")
  
  # Assign raster values of 0 to NA for cellSize calculation
  sst_depth_NA <- sst_depth_condition
  sst_depth_NA[sst_depth_NA == 0] <- NA
  
  # Find suitable area
  suitable_area <- cellSize(sst_depth_NA,
                            mask = TRUE,
                            unit = "km")
  
  
  # Rasterize the eez data
  eez_raster <- rasterize(eez, suitable_area, "rgn")
  
  # Sum suitable areas for each eez 
  eez_suitable <- zonal(x = suitable_area, 
                        z = eez_raster, 
                        fun = "sum", 
                        na.rm = TRUE)
  
  
  # Create a table of suitable area for the species in each eez 
  suitable_area_table <- kableExtra::kable(eez_suitable, col.names = c("Region", "Suitable Area (km\u00B2)"), align = "c")
  
  # Join eez_suitable data with the original eez data to map
  eez_join <- left_join(eez, eez_suitable, by = "rgn")
  
  
  # Map of suitable areas 
  map <- tm_shape(depth_crop) +
    tm_raster(palette = paletteer_c("ggthemes::Green-Blue Diverging", 7, direction = -1),
              title = "Depth (m)",
              midpoint = NA) +
    tm_shape(eez_join) +
    tm_polygons(col = "area",
                palette = paletteer_c("ggthemes::Classic Area-Brown", 5),
                title = "Suitable area (km\u00B2)",
                alpha = 0.9) +
    tm_text("rgn", 
            size = 0.45,
            col = "black") + 
    tm_layout(frame = FALSE,
              legend.outside = TRUE,
              legend.text.size = 0.65,
              main.title = paste("Suitable aquaculture area (km\u00B2) for", species, "along \nthe West Coast Exclusive Economic Zone (EEZ)"),
              main.title.size = 1.00) +
    tm_compass(position = c(0, 0.125),
               size = 1.5) +
    tm_scale_bar(position = c(0.05, 0.025)) 
  
  # Print both outputs, Kable table and Map 
  return(suitable_area_table)
  
  print(map)
}
```

```{r}
# Use my function for the California mussel, *Mytilus californianu*
# species = "California mussel"
# max_sst = 22
# min_sst = 5
# max_depth = 100,
# min_depth = 0 
aqua_fun("California", 22, 5, 100, 0)
```


## Citations
```{r}
#| code-fold: true
#| code-summary: "Reveal code"


tribble(
  ~Data, ~Citations, ~Link,
  "A review of California mussel (Mytilus californianus) fisheries biology and fisheries programs.", "	Schmidt, D. 1999 A review of California mussel (Mytilus californianus) fisheries biology and fisheries programs. Mytilus californianus; biology; Canadian Stock Assessment Secretariat Research Document 99/187.", "[Species](https://www.sealifebase.ca/summary/Mytilus-californianus.html)",
  "Mapping the global potential for marine aquaculture", "Gentry, R.R., Froehlich, H.E., Grimm, D. et al. Mapping the global potential for marine aquaculture. Nat Ecol Evol 1, 1317–1324 (2017). https://doi.org/10.1038/s41559-017-0257-9", "[Aqua](https://www.nature.com/articles/s41559-017-0257-9#citeas)"
) %>% 
  kableExtra::kable()

```


